### 2022-04-27 paper/localpsf.tex
add ellipsoid intersection figure
k=7

@@

If $\Aop$ is symmetric positive semidefinite, we symmetrize $\mathbf{A}_H$ as follows
\begin{equation*}
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right),
\end{equation*}
then we use the high order rational function spectral projection method from \cite{HackbuschKress07} to replace (or approximately replace) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero. In detail, let $f$ be the following scalar rational function,
\begin{equation*}
	f(\lambda) := \begin{cases}
		\lambda, & \lambda \ge 0, \\
		0, & \lambda < 0,
	\end{cases}
\end{equation*}
and extend $f$ to matrices via functional calculus. The matrix $f(\mathbf{A}_H^{\text{sym}})$ is the ideal positive semidefinite modification of $\mathbf{A}_H^{\text{sym}}$, because $f(\mathbf{A}_H^{\text{sym}})$ is the same as $\mathbf{A}_H^{\text{sym}}$ except all the negative eigenvalues are replaced with zero. Computing $f(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $f$ by the high order rational function
\begin{equation*}
	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - M}{M}\right)^{2^k} },
\end{equation*}
where $k$ is a positive integer, and $M$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$. In practice we use $k=5$, and $M=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration. If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (0,M)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[0,M]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$. This is illustrated in Figure \ref{fig:spd_rat}. 
\begin{figure}
	\centering
	\includegraphics[scale=0.5]{spd_rational_function.pdf}
	\caption{Rational function $\widetilde{f}(\lambda)$ for $M=1$ and $k=3,5,7$. As $k$ increases, $\widetilde{f}(\lambda)$ approaches the sawtooth function supported on $(0,M)$. }
	\label{fig:spd_rat}
\end{figure}

Using $\widetilde{f}$ instead of $f$ permits us to efficiently compute $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ using H-matrix methods. Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\widetilde{f}$ to matrices becomes
\begin{equation}
\label{eq:rational343}
\widetilde{f}(\mathbf{A}_H^{\text{sym}}) = \left(\mathbf{I} + \mathbf{X}^{2^k}\right)^{-1} \mathbf{A}_H^{\text{sym}},
\end{equation}
where
\begin{equation*}
	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - M \mathbf{I}\right)/M.
\end{equation*}
We compute $\mathbf{X}$ using fast H-matrix scalar multiplication and addition. Then we form the $\mathbf{X}^{2^k}$ via repeated squaring,
\begin{align*}
	\mathbf{X}^2 &= \mathbf{X} \mathbf{X} \\
	\mathbf{X}^4 &= \mathbf{X}^2 \mathbf{X}^2 \\
	&\vdots \\
	\mathbf{X}^{2^k} &= \mathbf{X}^{2^{k-1}} \mathbf{X}^{2^{k-1}},
\end{align*}
where each of these matrix squarings is performed using fast H-matrix multiplication. Finally, $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{X}^{2^k}$ and $\mathbf{A}_H^{\text{sym}}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition, matrix inversion, and matrix-matrix multiplication. At the end, the matrix is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic. Once constructed, we use the symmetric positive semidefinite matrix $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ in place of $\mathbf{A}_H$ in further linear algebra operations.

>>>>

\begin{figure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.40]{ellipsoids_intersect0.pdf}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.40]{ellipsoids_intersect1.pdf}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.40]{ellipsoids_intersect2.pdf}
	\end{subfigure}
	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
	}
	\label{fig:ellipsoid_intersection_test}
\end{figure}

@@

If $\Aop$ is symmetric positive semidefinite, we symmetrize $\mathbf{A}_H$ as follows
\begin{equation*}
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right),
\end{equation*}
then we use the high order rational function spectral projection method from \cite{HackbuschKress07} to replace (or approximately replace) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero. In detail, let $f$ be the following scalar rational function,
\begin{equation*}
	f(\lambda) := \begin{cases}
		\lambda, & \lambda \ge 0, \\
		0, & \lambda < 0,
	\end{cases}
\end{equation*}
and extend $f$ to matrices via functional calculus. The matrix $f(\mathbf{A}_H^{\text{sym}})$ is the ideal positive semidefinite modification of $\mathbf{A}_H^{\text{sym}}$, because $f(\mathbf{A}_H^{\text{sym}})$ is the same as $\mathbf{A}_H^{\text{sym}}$ except all the negative eigenvalues are replaced with zero. Computing $f(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $f$ by the high order rational function
\begin{equation*}
	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - (b+a)}{b-a}\right)^{2^k} },
\end{equation*}
where $k$ is a positive integer, $a$ is either zero, or a small positive real number, and $b$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$. If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (a,b)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[a,b]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$. This is illustrated in Figure \ref{fig:spd_rat}.  In practice we use $k=7$, $a=0$ and $b=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration. 
\begin{figure}
	\centering
	\includegraphics[scale=0.5]{spd_rational_function.pdf}
	\caption{Rational function $\widetilde{f}(\lambda)$ for $M=1$ and $k=3,5,7$. As $k$ increases, $\widetilde{f}(\lambda)$ approaches the sawtooth function supported on $(0,M)$. }
	\label{fig:spd_rat}
\end{figure}

Using $\widetilde{f}$ instead of $f$ permits us to efficiently compute $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ using H-matrix methods. Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\widetilde{f}$ to matrices becomes
\begin{equation}
\label{eq:rational343}
\widetilde{f}(\mathbf{A}_H^{\text{sym}}) = \left(\mathbf{I} + \mathbf{X}^{2^k}\right)^{-1} \mathbf{A}_H^{\text{sym}},
\end{equation}
where
\begin{equation*}
	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
\end{equation*}
We compute $\mathbf{X}$ using fast H-matrix scalar multiplication and addition. Then we form the $\mathbf{X}^{2^k}$ via repeated squaring,
\begin{align*}
	\mathbf{X}^2 &= \mathbf{X} \mathbf{X} \\
	\mathbf{X}^4 &= \mathbf{X}^2 \mathbf{X}^2 \\
	&\vdots \\
	\mathbf{X}^{2^k} &= \mathbf{X}^{2^{k-1}} \mathbf{X}^{2^{k-1}},
\end{align*}
where each of these matrix squarings is performed using fast H-matrix multiplication. Finally, $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{X}^{2^k}$ and $\mathbf{A}_H^{\text{sym}}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition, matrix inversion, and matrix-matrix multiplication. At the end, the matrix is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic. Once constructed, we use the symmetric positive semidefinite matrix $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ in place of $\mathbf{A}_H$ in further linear algebra operations.
